Apparent preferences for cognitive effort fade when multiple forms of effort

### 2020-12-08 Cost_manuscript.Rmd
first draft! (just add a fig)

  Understanding the sporadic attractiveness of effort remains elusive. In this study we sought to identify factors underlying this phenomenon, testing the hypothesis that the perceived cost or value of a demand could depend on the ability to achieve the same outcome through different means. Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found that preferences for cognitive effort were stable when individuals faced a single form of demand, but faded over time when other costs were also experienced.

@@

  These choice dynamics ruled out a number of common explanatory candidates. For instance, differences in choice stability between experiments suggested that a demand's cost is not intrinsic. The gradual choice convergence in the second experiment also discarded the effects of boredom, which would have accentuated the appeal of active tasks over time, or the possibility that our sample was mostly composed of demand seekers [@Cacioppo1996; @Sayali2019]. Even though fatigue could explain the decrease in effort acceptances across experiments, we would have expected tired participants to turn to wait trials over time in order to preserve earning rates. We also ruled out the sunk cost effects used to explain overharvesting [@Wikenheiser2013; @Sweis2018], as participants expressed their choices early and rarely quit during the handling time. Risk was also an unlikely factor, as task performance improved over time. Finally, the observed differences in acceptances between effort types could not be explained by reward alone [@Eisenberger1996], as demands were rewarded at the same rate.

@@

  Relying on foraging paradigms helped us to control for common shortfalls in the existing literature. For example, even though engaging with a demanding offer encompasses evaluating, pursuing, and receiving its outcome, investigations of the pursuit and receipt stages have been limited. Many studies have sought to isolate the evaluation stage by presenting individuals with offers that do not come into effect until the end of the experimental session (with delays spanning up to months), deliberately setting aside the question of how the ongoing experience of cost and outcome affects subsequent choices [@Kable2007; @Chong2016; @Massar2015; @Apps2015a]. However, discrepancies between expected and experienced value are well known [@Kahneman2006], and choosing immediate outcomes could be due to factors other than time preference, such as the risk of being unable to redeem the future rewards, or hoping to find better offers in the near future [@Fawcett2012; @Frederick2002; @Sozou1998]. We circumvented these pitfalls by providing participants with an opportunity to actively experience the demands they accepted, and gaving them full knowledge of the available options. Another challenge in comparing effort and delay is that increments in the required effort are often paired with a longer execution time. This issue is often avoided by comparing discounting functions, which yield points at which individuals faced with increasing costs become indifferent to large rewards [@Massar2015; @Prevost2010]. We eliminated this possibility by fixing the handling time for all demands, allowing us to probe people's preferences directly. Moreover, the between-subjects experiment, which resembled the traditional way costs are compared, provided a baseline of foraging behavior that contextualized the findings of the second experiment (a direct test of our hypothesis). Finally, the existence of optimal strategies helped us identify when demands could be attractive, a central motivation for this study. The apparent attractiveness of cognitive effort was reflected both by a higher proportion of acceptances at the expense of earnings, and by comparisons of fitted opportunity costs against the optimal earning rate. Of note is that the optimality of the strategies induced by each demand depended on the richness of the block. This can help explain why the wait group earned the most despite having similar acceptance rates to the physical group, as the former was closer to optimality in the richest environment.

@@

  The present findings have important implications for daily and clinical scenarios. Excessive aversion to delay and effort can manifest as impulsivity and apathy, respectively [@Ainslie1975; @Bonnelle2016; @Chong2016; @Cummings1993], fluctuations in which are hallmark symptoms of various mental illnesses and substance use disorders [@Paulus2007]. Lack of motivation to exert effort or wait for prospects is a distinguishing factor among mental afflictions like depression and schizophrenia [@Barch2019]. Within the former, medicated patients in remission are more sensitive to effort than controls, and this sensitivity is predictive of relapse [@Berwian2020]. Encouraging patients to engage in activities has been found to be a helpful treatment for depression [@Dimidjian2006]. Tailored reward schedules can successfully achieve this [@Eisenberger1996]. However, since depressed patients also show insensitivity to rewards [@Huys2013], identifying what makes effort appealing in its own right can help breach treatment-resistant cases. Our work suggests that subdividing a task can promote its pursual either by taxing working memory and/or allowing individuals to count completed tasks instead of continuous time. More importantly, it highlights that decision scenarios should be configured to promote pursual of desired activities, perhaps by removing alternative uses of time. These same principles can potentially be applied to increasing productivity at work or adherence to physical exercise routines.

### *Points to cover*

- Regardless, our focus here was not on the amount of effort exherted, but on what makes people pursue the task in spite of its toll.

- we have all felt the need to populate empty time

- It's maybe relevant that waiting is goal oriented (you don't sit around for its own sake), but activities are pursued for their own appeal. *like, waiting is a goal oriented activity, but effort can be the goal itself.* in this sense it makes sense to study delays in terms of their modulation of the value of prospects, but we have unnecessarily adopted this view for ethe study of ffort, thinking of it mostly as a cost. Even if their discounting properties are similar, their resulting behavior can be contrasting (apathy vs impulsivity), and they're impacted differently by mental illness (adhd amygdala paper). Also, it's not just that we would rather to fill in empty periods of waiting, but how.

- *Theories enumerated by Eisenberger and Cameron (1996) focus on how people see their behavior to explain why tasks are intrinsically attractive. We, on the other hand, point to task-specific features that promote appeal, which are amenable to configuration in everyday decision contexts. In other words, self-explanatory motivation versus task-specific ones.*

>>>>

  Understanding the sporadic attractiveness of effort remains elusive. In this study, we sought to identify factors underlying this phenomenon, and tested the hypothesis that the perceived cost or value of a demand could depend on the ability to achieve the same outcome through different means. Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found that preferences for cognitive effort were stable when individuals faced a single form of demand, but faded over time when other costs were also experienced.

@@

  These choice dynamics ruled out a number of common explanatory candidates. For instance, differences in choice stability between experiments suggested that a demand's cost is not intrinsic. In addition, the gradual choice convergence in the second experiment also discarded the effects of boredom, which would have accentuated the appeal of active tasks over time, or the possibility that our sample was mostly composed of demand seekers [@Cacioppo1996; @Sayali2019]. Even though fatigue could explain the decrease in effort acceptances across experiments, we would have expected tired participants to turn to wait trials over time in order to preserve earning rates. We also ruled out the sunk cost effects used to explain overharvesting [@Wikenheiser2013; @Sweis2018], as participants expressed their choices early and rarely quit during the handling time. Risk was also an unlikely factor, as task performance improved over time. Finally, the observed differences in acceptances between effort types could not be explained by reward alone [@Eisenberger1996], as demands were rewarded at the same rate.

@@

  Relying on foraging paradigms helped us to control for common shortfalls in the existing literature. For example, even though engaging with a demanding offer encompasses evaluating, pursuing, and receiving its outcome, investigations of the pursuit and receipt stages have been limited. Many studies have sought to isolate the evaluation stage by presenting individuals with offers that do not come into effect until the end of the experimental session (with delays spanning up to months), deliberately setting aside the question of how the ongoing experience of cost and outcome affects subsequent choices [@Kable2007; @Chong2016; @Massar2015; @Apps2015a]. However, discrepancies between expected and experienced value are well known [@Kahneman2006], and choosing immediate outcomes could be due to factors other than time preference, such as the risk of being unable to redeem the future rewards, or hoping to find better offers in the near future [@Fawcett2012; @Frederick2002; @Sozou1998]. We circumvented these pitfalls by providing participants with an opportunity to actively experience the demands they accepted, and gaving them full knowledge of the available options. Another challenge in comparing effort and delay is that increments in the required effort are often paired with a longer execution time. This issue is often avoided by comparing discounting functions, which yield points at which individuals faced with increasing costs become indifferent to large rewards [@Massar2015; @Prevost2010]. We eliminated this possibility by fixing the handling time for all demands, allowing us to probe people's preferences directly. Moreover, the between-subjects experiment, which resembled the traditional way costs are compared, provided a baseline of foraging behavior that contextualized the findings of the second experiment (a direct test of our hypothesis). Finally, the existence of optimal strategies helped us identify when demands could be attractive, a central motivation for this study. The apparent attractiveness of cognitive effort was reflected both by a higher proportion of acceptances at the expense of earnings, and by the fitted opportunity costs biases discussed above. Of note is that the optimality of the strategies induced by each demand depended on the richness of the block. This can help explain why the wait group earned the most despite having similar acceptance rates to the physical group, as the former was closer to optimality in the richest environment.

@@

  The present findings have important implications for daily and clinical scenarios. Excessive aversion to delay and effort can manifest as impulsivity and apathy, respectively [@Ainslie1975; @Bonnelle2016; @Chong2016; @Cummings1993], fluctuations in which are hallmark symptoms of various mental illnesses and substance use disorders [@Paulus2007]. Lack of motivation to exert effort or wait for prospects is a distinguishing factor among mental afflictions like depression and schizophrenia [@Barch2019]. Within the former, medicated patients in remission are more sensitive to effort than controls, and this sensitivity is predictive of relapse [@Berwian2020]. Encouraging patients to engage in activities has been found to be a helpful treatment for depression [@Dimidjian2006]. Tailored reward schedules can successfully achieve this [@Eisenberger1996], but are limited by the attenuated sensitivity to rewards in depressed patients. However, since depressed patients also show insensitivity to rewards [@Huys2013], identifying what makes effort appealing in its own right can help breach treatment-resistant cases. Our work suggests that subdividing a task can promote its pursual, although further testing is required. More importantly, it highlights that decision scenarios should be configured to promote pursual of desired activities, perhaps by removing alternative uses of time. These same principles can potentially be applied to increasing productivity at work or adherence to physical exercise routines.

### Conclusion

  In this study, we have examined the factors that motivate the preference of different demands, emphasizing how these can be molded by the decision maker's environment. Unlike existing theories that have focused on the role of incentives and retrospectives [@Eisenberger1996; @Inzlicht2018], our work highlights outcome-independent features that can motivate the pursual of effortful avenues, and which can be potentially manipulated in order to promote engagement in clinical and daily activities.
